---
title: "R Notebook"
output: html_notebook
---


library(tidyverse)
library(censusapi)
library(dplyr)

```{r}
#created a text file with the api key, read in
key=read_lines('data/key.txt')
```
```{r}
#create dataset for us census apis
apis <- listCensusApis()
#view(apis)
```
```{r}
#look into metadata for acs1/2019
metadata <- listCensusMetadata(
  name = "acs/acs1/profile",
  vintage = 2019)
```


```{r}
#view(metadata)
```
```{r}
#listCensusMetadata(type = "variables")
```

Pull in age demographics for 18 and above, then combine into decade buckets, and create a % for each bucket
```{r}

# create an age variable to plug into code, note we can only find estimates and not percentages for age & sex

age_vars <- c(
#Male_15_to_17_years = "B01001_006E",
#Female_15_to_17_years = "B01001_030E",
Male_18_and_19_years = "B01001_007E",
Male_20_years = "B01001_008E",
Female_18_and_19_years = "B01001_031E",
Female_20_years = "B01001_032E",
Male_21_years = "B01001_009E",
Male_22_to_24_years = "B01001_010E",
Male_25_to_29_years = "B01001_011E",
Female_21_years = "B01001_033E",
Female_22_to_24_years = "B01001_034E",
Female_25_to_29_years = "B01001_035E",
Male_30_to_34_years = "B01001_012E",
Male_35_to_39_years = "B01001_013E",
Female_30_to_34_years = "B01001_036E",
Female_35_to_39_years = "B01001_037E",
Male_40_to_44_years = "B01001_014E",
Male_45_to_49_years = "B01001_015E",
Female_40_to_44_years = "B01001_038E",
Female_45_to_49_years = "B01001_039E",
Male_50_to_54_years = "B01001_016E",
Male_55_to_59_years = "B01001_017E",
Female_50_to_54_years = "B01001_040E",
Female_55_to_59_years = "B01001_041E",
Male_60_and_61_years = "B01001_018E",
Male_62_to_64_years = "B01001_019E",
Male_65_and_66_years = "B01001_020E",
Female_60_and_61_years = "B01001_042E",
Female_62_to_64_years = "B01001_043E",
Female_65_and_66_years = "B01001_044E",
Male_67_to_69_years = "B01001_021E",
Male_70_to_74_years = "B01001_022E",
Male_75_to_79_years = "B01001_023E",
Male_80_to_84_years = "B01001_024E",
Male_85_years_and_over = "B01001_025E",
Female_67_to_69_years = "B01001_045E",
Female_70_to_74_years = "B01001_046E",
Female_75_to_79_years = "B01001_047E",
Female_80_to_84_years = "B01001_048E",
Female_85_years_and_over = "B01001_049E")


```
plus age variable into code below to pull in data for Davidson County
```{r}
state_age <- map_dfr(
    years,
    ~ get_acs(
        geography = "county",
        variables = c(age_vars),
        year = .x,
        survey = 'acs1'
    ),
    .id = 'year'
    )  %>% 
    clean_names()
state_age <- state_age %>% filter(geoid=='47037')

#view(metadata)
```

Aggregate the age groups into decade buckets (e.g., 20-29, 30-39) 
```{r}
```
```{r}
state_age1 <- state_age %>% mutate(label = case_when(
  variable %in% c("B01001_007")~ "Male - 18 to 19 years",
  variable %in% c("B01001_008", "B01001_009", "B01001_010", "B01001_011")~ "Male - 20 to 29 years",
  variable %in% c("B01001_012", "B01001_013")~ "Male - 30 to 39 years",
  variable %in% c("B01001_014", "B01001_015")~ "Male - 40 to 49 years",
  variable %in% c("B01001_016", "B01001_017")~ "Male - 50 to 59 years",
  variable %in% c("B01001_018", "B01001_019", "B01001_020")~ "Male - 60 to 66 years",
  variable %in% c("B01001_021", "B01001_022", "B01001_023", "B01001_024", "B01001_025")~ "Male - 67 and over", 
  variable %in% c("B01001_031")~ "Female - 18 to 19 years",
  variable %in% c("B01001_032", "B01001_033", "B01001_034", "B01001_035")~ "Female - 20 to 29 years",
variable %in% c("B01001_036", "B01001_037") ~ "Female - 30 to 39 years",
variable %in% c("B01001_038", "B01001_039") ~ "Female - 40 to 49 years",
variable %in% c("B01001_040", "B01001_041") ~ "Female - 50 to 59 years",
variable %in% c("B01001_042", "B01001_043", "B01001_044") ~ "Female - 60 to 66 years",
variable %in% c("B01001_045", "B01001_046", "B01001_047", "B01001_048", "B01001_049") ~ "Female - 67 and over"))


```
convert each bucket to percentage 
```{r}
state_age2 <- state_age1%>% 
  group_by(label) %>% 
  summarise(estimate = sum(estimate)) %>% 
  mutate(percent_age = estimate/sum(estimate) * 100)

#view(state_age2)


```


```{r}
age_bracket <- state_age2[,-c(2)]
age_bracket <- age_bracket %>% 
  rename(Variables = label, Percentage = percent_age)
view(age_bracket)
 
```

```{r}
#pull in variables for race, gender, age, education by percentage. Exclude #disability, veteran, sexual orientation (diff data source for sex orient)

acs1 <- getCensus(
  key = key,
  name = "acs/acs1/profile", 
  vintage = 2019,

  vars = c("DP05_0033PE", "DP05_0039PE", "DP05_0038PE", "DP05_0044PE", "DP05_0052PE","DP05_0037PE", "DP05_0035PE", "DP05_0027PE", "DP05_0026PE", "DP05_0071PE", "DP02_0060PE", "DP02_0067PE", "DP02_0063PE", "DP02_0064PE", "DP02_0065PE", "DP02_0066PE", "DP05_0057PE"),
  #Davidson County
  region = 'county:037', 
  #State of TN
  regionin = 'state:47') 

```
Rename columns
```{r}
acs1 <- acs1 %>% 
  rename("Total Pop" = DP05_0033PE, TN = state, Davidson = county, "American Indian and Alaska Native" = DP05_0039PE, Asian = DP05_0044PE, "Black or African American" = DP05_0038PE, "Native Hawaiian and Other Pacific Islander" = DP05_0052PE, "White Alone" = DP05_0037PE, "Hispanic or Latino" = DP05_0071PE, "Two or more races" = DP05_0035PE, "Male - 18 and over" = DP05_0026PE, "Female - 18 and over" = DP05_0027PE, "Less than 9th Grade" = DP02_0060PE, "High school graduate (includes equivalency)" = DP02_0067PE, "Some college, no degree" = DP02_0063PE, "Associate's degree" = DP02_0064PE, "Bachelor's degree" = DP02_0065PE, "Graduate/Professional Degree" = DP02_0066PE, "Other Race" = DP05_0057PE)
```
(Placeholder for civilian_vets = DP02_0070PE, disabled_18_to_64 = DP02_0076PE, total_pop_not_hispanic_lat = DP05_0076PE in case needed later)

```{r}
#convert to test csv file
#write.csv(acs1,"data/test_file.csv", row.names = TRUE)
```
change acs1 file from horizontal to vertical layout

```{r}
acs1_long <- gather(acs1, Variables, Percentage)

```
```{r}
acs1_vars <- acs1_long[-c(1,2,3),]
```
```{r}
view(acs1_vars)
```
Append acs1_vars with age_bracket and call it data_comb
First, determine column types for conformity
```{r}
list(acs1_vars, age_bracket)
```
```{r}
#convert the Percentage column on both datasets to numeric
acs1_num <- acs1_vars %>% 
  mutate_at(c(2), as.numeric)

```

```{r}
# for age_bracket, shorten the Percentage column to 1 decimal point
#age_bracket <- age_bracket %>% 
#~ round(., 0)
  
  
```


```{r}
data_comb <- bind_rows(acs1_num, age_bracket)
data_comb
```
```{r}
#convert to test csv file
write.csv(data_comb,"data/test_file2.csv", row.names = TRUE)
```

