---
title: "dq5 People3"
author: "Alvin Wendt"
date: "11/24/2020"
output: html_document
---
Libraries to import
```{r}
library(tidycensus)
library(censusapi)
library(janitor)
#library(tigris)
#options(tigris_use_cache = TRUE)
library(tidyverse)
library(purrr)
```
Apply API Key to Variable
```{r}
api_key <- "input API Key here"
```

Install Census API key
```{r}
census_api_key(api_key,install = TRUE)
```
Load ACS Variable list
```{r}
acs_variable_list <- load_variables(2019,"acs1", cache = TRUE)
```

ACS Variable List write to CSV for easy viewing and filtering
```{r}
write.csv(acs_variable_list,'acs_variable_list_2019',row.names = FALSE)
```

Code to run Metadata information in determining variable
```{r}
key=read_lines(api_key)

apis <- listCensusApis()
view(apis)

metadata <- listCensusMetadata(
    name = 'acs/acs1/profile',
    vintage = 2019)

```

Create ACS Race Variables
```{r}
race_variables_pecentage <- c(totpop = 'DP05_0033PE',
                     nothispanic = 'DP05_0076PE',
                     nh_whitealone = 'DP05_0037PE',
                     nh_black = 'DP05_0038PE',
                     nh_indian = 'DP05_0039PE',
                     nh_asian = 'DP05_0044PE',
                     nh_pacisl = 'DP05_0052PE',
                     nh_multi = 'DP05_0035PE',
                     hispanic = 'DP05_0071PE')
```

Create Education
```{r}
education_variables <- c(Pop25andover = 'DP02_0059E',
                         lessthan9th = 'DP02_0060PE',
                         highschool = 'DP02_0067PE',
                         somecollege = 'DP02_0063PE',
                         associates = 'DP02_0064PE',
                         bachelors = 'DP02_0065PE',
                         graduate = 'DP02_0066PE')
```

Create Age Variables
```{r}
age_variables <- c(SEX_BY_AGE_Total = "B01001_001E",
SEX_BY_AGE_Male_ = "B01001_002E",
SEX_BY_AGE_Male_Under_5_years = "B01001_003E",
SEX_BY_AGE_Male_5_to_9_years = "B01001_004E",
SEX_BY_AGE_Male_10_to_14_years = "B01001_005E",
SEX_BY_AGE_Male_15_to_17_years = "B01001_006E",
SEX_BY_AGE_Male_18_and_19_years = "B01001_007E",
SEX_BY_AGE_Male_20_years = "B01001_008E",
SEX_BY_AGE_Male_21_years = "B01001_009E",
SEX_BY_AGE_Male_22_to_24_years = "B01001_010E",
SEX_BY_AGE_Male_25_to_29_years = "B01001_011E",
SEX_BY_AGE_Male_30_to_34_years = "B01001_012E",
SEX_BY_AGE_Male_35_to_39_years = "B01001_013E",
SEX_BY_AGE_Male_40_to_44_years = "B01001_014E",
SEX_BY_AGE_Male_45_to_49_years = "B01001_015E",
SEX_BY_AGE_Male_50_to_54_years = "B01001_016E",
SEX_BY_AGE_Male_55_to_59_years = "B01001_017E",
SEX_BY_AGE_Male_60_and_61_years = "B01001_018E",
SEX_BY_AGE_Male_62_to_64_years = "B01001_019E",
SEX_BY_AGE_Male_65_and_66_years = "B01001_020E",
SEX_BY_AGE_Male_67_to_69_years = "B01001_021E",
SEX_BY_AGE_Male_70_to_74_years = "B01001_022E",
SEX_BY_AGE_Male_75_to_79_years = "B01001_023E",
SEX_BY_AGE_Male_80_to_84_years = "B01001_024E",
SEX_BY_AGE_Male_85_years_and_over = "B01001_025E",
SEX_BY_AGE_Female_ = "B01001_026E",
SEX_BY_AGE_Female_Under_5_years = "B01001_027E",
SEX_BY_AGE_Female_5_to_9_years = "B01001_028E",
SEX_BY_AGE_Female_10_to_14_years = "B01001_029E",
SEX_BY_AGE_Female_15_to_17_years = "B01001_030E",
SEX_BY_AGE_Female_18_and_19_years = "B01001_031E",
SEX_BY_AGE_Female_20_years = "B01001_032E",
SEX_BY_AGE_Female_21_years = "B01001_033E",
SEX_BY_AGE_Female_22_to_24_years = "B01001_034E",
SEX_BY_AGE_Female_25_to_29_years = "B01001_035E",
SEX_BY_AGE_Female_30_to_34_years = "B01001_036E",
SEX_BY_AGE_Female_35_to_39_years = "B01001_037E",
SEX_BY_AGE_Female_40_to_44_years = "B01001_038E",
SEX_BY_AGE_Female_45_to_49_years = "B01001_039E",
SEX_BY_AGE_Female_50_to_54_years = "B01001_040E",
SEX_BY_AGE_Female_55_to_59_years = "B01001_041E",
SEX_BY_AGE_Female_60_and_61_years = "B01001_042E",
SEX_BY_AGE_Female_62_to_64_years = "B01001_043E",
SEX_BY_AGE_Female_65_and_66_years = "B01001_044E",
SEX_BY_AGE_Female_67_to_69_years = "B01001_045E",
SEX_BY_AGE_Female_70_to_74_years = "B01001_046E",
SEX_BY_AGE_Female_75_to_79_years = "B01001_047E",
SEX_BY_AGE_Female_80_to_84_years = "B01001_048E",
SEX_BY_AGE_Female_85_years_and_over = "B01001_049E"
)
```

Create Veteran/Disabled Filter Variables
```{r}
vet_disabled_variables <- c(civilian_vets = 'DP02_0070PE',
                         disabled_18_to_64 = 'DP02_0076PE'
                         )
```

Create Years Filter Variables
```{r}
years <- lst(2019)
```

Run Davidson County table for Race variables
```{r}
state_diversity <- map_dfr(
    years,
    ~ get_acs(
        geography = "county",
        variables = race_variables_pecentage,
        year = .x,
        survey = 'acs1'
    ),
    .id = 'year'
    ) %>% clean_names()

state_diversity <- state_diversity %>% filter(geoid=='47037')
```

Create Column for Race Labels
```{r}
state_diversity$Label[state_diversity$variable ==  "DP05_0033P"] <- "totpop"
state_diversity$Label[state_diversity$variable ==  "DP05_0076P"] <- "nothispanic"
state_diversity$Label[state_diversity$variable ==  "DP05_0037P"] <- "nh_whitealone"
state_diversity$Label[state_diversity$variable ==  "DP05_0038P"] <- "nh_black"
state_diversity$Label[state_diversity$variable ==  "DP05_0039P"] <- "nh_indian"
state_diversity$Label[state_diversity$variable ==  "DP05_0044P"] <- "nh_asian"
state_diversity$Label[state_diversity$variable ==  "DP05_0052P"] <- "nh_pacisl"
state_diversity$Label[state_diversity$variable ==  "DP05_0035P"] <- "nh_multi"
state_diversity$Label[state_diversity$variable ==  "DP05_0071P"] <- "hispanic"
```

Run Davidson County table for Education variables
```{r}
state_education <- map_dfr(
    years,
    ~ get_acs(
        geography = "county",
        variables = education_variables,
        year = .x,
        survey = 'acs1'
    ),
    .id = 'year'
    ) %>% clean_names()

state_education %>% filter(geoid=='47037')
```

Create Column for Education Labels
```{r}
state_education$Label[state_education$variable ==  "DP02_0059"] <- "Pop25andover"
state_education$Label[state_education$variable ==  "DP02_0060P"] <- "lessthan9th"
state_education$Label[state_education$variable ==  "DP02_0067P"] <- "highschool"
state_education$Label[state_education$variable ==  "DP02_0063P"] <- "somecollege"
state_education$Label[state_education$variable ==  "DP02_0064P"] <- "associates"
state_education$Label[state_education$variable ==  "DP02_0065P"] <- "bachelors"
state_education$Label[state_education$variable ==  "DP02_0066P"] <- "graduate"


```


Run Davidson County table for Age variable
```{r}
state_age <- map_dfr(
    years,
    ~ get_acs(
        geography = "county",
        variables = age_variables,
        year = .x,
        survey = 'acs1'
    ),
    .id = 'year'
    )  %>%
    clean_names()

state_age <- state_age %>% filter(geoid=='47037')
```
Create Column for Age Labels
```{r}
state_age$Label[state_age$variable == "B01001_001"] <- "SEX_BY_AGE_Total"
state_age$Label[state_age$variable == "B01001_002"] <- "SEX_BY_AGE_Male_"
state_age$Label[state_age$variable == "B01001_003"] <- "SEX_BY_AGE_Male_Under_5_years"
state_age$Label[state_age$variable == "B01001_004"] <- "SEX_BY_AGE_Male_5_to_9_years"
state_age$Label[state_age$variable == "B01001_005"] <- "SEX_BY_AGE_Male_10_to_14_years"
state_age$Label[state_age$variable == "B01001_006"] <- "SEX_BY_AGE_Male_15_to_17_years"
state_age$Label[state_age$variable == "B01001_007"] <- "SEX_BY_AGE_Male_18_and_19_years"
state_age$Label[state_age$variable == "B01001_008"] <- "SEX_BY_AGE_Male_20_years"
state_age$Label[state_age$variable == "B01001_009"] <- "SEX_BY_AGE_Male_21_years"
state_age$Label[state_age$variable == "B01001_010"] <- "SEX_BY_AGE_Male_22_to_24_years"
state_age$Label[state_age$variable == "B01001_011"] <- "SEX_BY_AGE_Male_25_to_29_years"
state_age$Label[state_age$variable == "B01001_012"] <- "SEX_BY_AGE_Male_30_to_34_years"
state_age$Label[state_age$variable == "B01001_013"] <- "SEX_BY_AGE_Male_35_to_39_years"
state_age$Label[state_age$variable == "B01001_014"] <- "SEX_BY_AGE_Male_40_to_44_years"
state_age$Label[state_age$variable == "B01001_015"] <- "SEX_BY_AGE_Male_45_to_49_years"
state_age$Label[state_age$variable == "B01001_016"] <- "SEX_BY_AGE_Male_50_to_54_years"
state_age$Label[state_age$variable == "B01001_017"] <- "SEX_BY_AGE_Male_55_to_59_years"
state_age$Label[state_age$variable == "B01001_018"] <- "SEX_BY_AGE_Male_60_and_61_years"
state_age$Label[state_age$variable == "B01001_019"] <- "SEX_BY_AGE_Male_62_to_64_years"
state_age$Label[state_age$variable == "B01001_020"] <- "SEX_BY_AGE_Male_65_and_66_years"
state_age$Label[state_age$variable == "B01001_021"] <- "SEX_BY_AGE_Male_67_to_69_years"
state_age$Label[state_age$variable == "B01001_022"] <- "SEX_BY_AGE_Male_70_to_74_years"
state_age$Label[state_age$variable == "B01001_023"] <- "SEX_BY_AGE_Male_75_to_79_years"
state_age$Label[state_age$variable == "B01001_024"] <- "SEX_BY_AGE_Male_80_to_84_years"
state_age$Label[state_age$variable == "B01001_025"] <- "SEX_BY_AGE_Male_85_years_and_over"
state_age$Label[state_age$variable == "B01001_026"] <- "SEX_BY_AGE_Female_"
state_age$Label[state_age$variable == "B01001_027"] <- "SEX_BY_AGE_Female_Under_5_years"
state_age$Label[state_age$variable == "B01001_028"] <- "SEX_BY_AGE_Female_5_to_9_years"
state_age$Label[state_age$variable == "B01001_029"] <- "SEX_BY_AGE_Female_10_to_14_years"
state_age$Label[state_age$variable == "B01001_030"] <- "SEX_BY_AGE_Female_15_to_17_years"
state_age$Label[state_age$variable == "B01001_031"] <- "SEX_BY_AGE_Female_18_and_19_years"
state_age$Label[state_age$variable == "B01001_032"] <- "SEX_BY_AGE_Female_20_years"
state_age$Label[state_age$variable == "B01001_033"] <- "SEX_BY_AGE_Female_21_years"
state_age$Label[state_age$variable == "B01001_034"] <- "SEX_BY_AGE_Female_22_to_24_years"
state_age$Label[state_age$variable == "B01001_035"] <- "SEX_BY_AGE_Female_25_to_29_years"
state_age$Label[state_age$variable == "B01001_036"] <- "SEX_BY_AGE_Female_30_to_34_years"
state_age$Label[state_age$variable == "B01001_037"] <- "SEX_BY_AGE_Female_35_to_39_years"
state_age$Label[state_age$variable == "B01001_038"] <- "SEX_BY_AGE_Female_40_to_44_years"
state_age$Label[state_age$variable == "B01001_039"] <- "SEX_BY_AGE_Female_45_to_49_years"
state_age$Label[state_age$variable == "B01001_040"] <- "SEX_BY_AGE_Female_50_to_54_years"
state_age$Label[state_age$variable == "B01001_041"] <- "SEX_BY_AGE_Female_55_to_59_years"
state_age$Label[state_age$variable == "B01001_042"] <- "SEX_BY_AGE_Female_60_and_61_years"
state_age$Label[state_age$variable == "B01001_043"] <- "SEX_BY_AGE_Female_62_to_64_years"
state_age$Label[state_age$variable == "B01001_044"] <- "SEX_BY_AGE_Female_65_and_66_years"
state_age$Label[state_age$variable == "B01001_045"] <- "SEX_BY_AGE_Female_67_to_69_years"
state_age$Label[state_age$variable == "B01001_046"] <- "SEX_BY_AGE_Female_70_to_74_years"
state_age$Label[state_age$variable == "B01001_047"] <- "SEX_BY_AGE_Female_75_to_79_years"
state_age$Label[state_age$variable == "B01001_048"] <- "SEX_BY_AGE_Female_80_to_84_years"
state_age$Label[state_age$variable == "B01001_049"] <- "SEX_BY_AGE_Female_85_years_and_over"


```

Run Davidson County table for Veteran/Disabled variables
```{r}
state_vet_disabled <- map_dfr(
    years,
    ~ get_acs(
        geography = "county",
        variables = vet_disabled_variables,
        year = .x,
        survey = 'acs1'
    ),
    .id = 'year'
    )  %>%
    clean_names()

state_vet_disabled <- state_vet_disabled %>% filter(geoid=='47037')
```
Create Column for Labels
```{r}
state_vet_disabled$Label[state_vet_disabled$variable == "DP02_0070P"] <- "civilian_vets"
state_vet_disabled$Label[state_vet_disabled$variable == "DP02_0076P"] <- "disabled_18_to_64"
```















